Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2022-10-04 paper-correlations.tex
Fix labels

\section{Background and motivation}
Materials are ubiquitous both in nature and in industrial human activities, and
their physical properties are interconnected to
structure\cite{Torquato_book}\cite{Sahimi_book}\cite{Adler_recon}. The information
about the structure usually comes in the way of 2D and 3D images obtained by
microscopy, tomography and other related methodologies. This structural
information can be used for numerous types of analysis, including simulations of
different processes within studied materials or evaluation of their properties
using so-called pore-scale modelling. Unlike computational modelling techniques
a whole class of approximate methods exists including (rigorous) bounds that
allows very fast estimations of physical properties albeit the bounds themselves
are too wide to be applicable in majority of practical problems. Nonetheless, it
is noteworthy that such bounds are originally based on structural descriptors in
the form of correlation functions.

@@

The aim of this paper is to establish a general open-source solution allowing to
compute all classical correlation functions. This solution should be well
documented with all computational algorithms explained, it should also leverage
on recent advances in programming high-intensity computations and utilize both
CPU and GPU power. Our answer in the form of \code{CorrelationFunctions.jl}
package allows to compute numerous CFs easily under any operating system. The
manuscript describing the package is organized as follows: in \cref{scansec} we
provide all algorithmic details for CFs computation in directions (scanning
method). \cref{mapsec} presents all algorithmic details for computing full CFs
maps (map method, full 2-point statistics for a given image). In \cref{verisec}
we verify the all computations based on known analytical
solutions. \cref{effisec} deals with computational efficiency of our code, we
compare computational times for 2D and 3D images of different sizes for scanning
and map methods as evaluated using either CPU or GPU. In \cref{exasec} some
application examples are provided, including CFs evaluation for multi-phase
materials. Finally, \cref{sumsec} summarizes all results and outlines possible
future improvements.

\section{Overview of supported functions and modules.}
\label{oversec}
\verb+CorrelationFunctions.jl+ is capable to compute the following correlation
functions:
\begin{itemize}
\item Lineal-path function $L_2$.
\item Two-point function $S_2$.
\item Cluster function $C_2$.
\item Surface-surface function $F_{ss}$.
\item Surface-void function $F_{sv}$.
\item Pore size function $P$.
\item Chord length function $p$.
\item Phase cross-correlation function $\rho_{ij}$
\end{itemize}

These correlation functions are calculated using two different approaches (when
applicable). The first approach is implemented in the module \code{Directional}
and the second approach is in the module \code{Map}. The former approach (called
the scan approach) calculates correlation functions along specified directions
which can include axial and diagonal directions. The results of computations can
be provided to the user separately for each requested direction or averaged
across multiple directions. The latter approach (called the map approach)
calculates a ``correlation map'' which holds the values of a correlation
function in all directions. The scan and map approaches are described in
\cref{scansec} and \cref{mapsec} respectively.

@@

\section{The scan approach.}
\label{scansec}
In this approach, the input array is cut in parallel slices (``scanned'') along
each requested direction. Each slice is then analyzed independently as
one-dimensional system and the result is accumulated into \code{CorrelationData}
structure. In the following sections we provide the algorithms for computation
of the functions given in \cref{oversec}.

@@

\subsection{Lineal-path function}
Lineal-path function $L_2^{(i)}(\bm{r})$ is defined as the probability that a
line segment with the length $|\bm{r}|$ and paralel to $\bm{r}$ lies wholly in
the phase $i$ when randomly thrown into the input array. This function for a
one-dimensional array is computed using the algorithm described in
\apref{linpathalg}. A multidimensional arrays is firstly cut in slices and
\code{l2} function is calculated for each slice returning \code{success} and
\code{total} arrays. These arrays are then accumulated as described in
\cref{scansec} for final calculation of $L_2^{(i)}(\bm{r})$ function.

@@

\section{The map approach.}
\label{mapsec}

@@

\section{Verifying our algorithms}
\label{verisec}
To verify our package we use random datasets of overlapping disks (the number of
dimensions $n = 2$) or balls ($n = 3$) with radii $R$ and centers placed in
points generated by Poisson point process with a parameter $\lambda$. For these
datasets correlation functions can be given in closed-form expressions for one-,
two- and three-dimensional cases. For the sake of testing we obtain a collection
of disks (resp. balls) in a square (resp. cube) with the side $s$ using the
following algorithm:
\begin{algorithmic}[1]
  \label{testdata}
  \Procedure{testdata}{$s, n, R, \lambda$}
    \State $nballs \gets poisson(\lambda s^n)$ \Comment{$poisson(\lambda)$
      generates random numbers with Poisson distribution with parameter
      $\lambda$.}
    \State Generate $nballs$ random points which are uniformly distributes in
    a cube ($n = 3$) or square ($n = 2$) with a side $s$
    \State Draw balls (disks) of radius $R$ in those points. Store the result in
    an array $output$.
    \State \textbf{return} $output$
  \EndProcedure
\end{algorithmic}

@@

\section{Efficiency of our algorithms}
\label{effisec}
To measure efficiecy of our algorithms we use a machine with the following
configuration:
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 5 1600X
\item \textbf{Memory}: 32~Gb DDR4-2400 memory
\item \textbf{GPU}: Nvidia Tesla V100 with 32~Gb video memory
\end{itemize}

@@

\section{Examples and applications}
\label{exasec}
After the verification of our computational approach based on analytical
solution we can apply our algorithms to evaluate correlation functions for
different porous media images, including real XCT and SEM images.

@@

\section{Summary}
\label{sumsec}
Text

>>>>

\section{Background and motivation}
\label{sec:background}
Materials are ubiquitous both in nature and in industrial human activities, and
their physical properties are interconnected to
structure\cite{Torquato_book}\cite{Sahimi_book}\cite{Adler_recon}. The information
about the structure usually comes in the way of 2D and 3D images obtained by
microscopy, tomography and other related methodologies. This structural
information can be used for numerous types of analysis, including simulations of
different processes within studied materials or evaluation of their properties
using so-called pore-scale modelling. Unlike computational modelling techniques
a whole class of approximate methods exists including (rigorous) bounds that
allows very fast estimations of physical properties albeit the bounds themselves
are too wide to be applicable in majority of practical problems. Nonetheless, it
is noteworthy that such bounds are originally based on structural descriptors in
the form of correlation functions.

@@

The aim of this paper is to establish a general open-source solution allowing to
compute all classical correlation functions. This solution should be well
documented with all computational algorithms explained, it should also leverage
on recent advances in programming high-intensity computations and utilize both
CPU and GPU power. Our answer in the form of \code{CorrelationFunctions.jl}
package allows to compute numerous CFs easily under any operating system. The
manuscript describing the package is organized as follows: in \cref{sec:scan} we
provide all algorithmic details for CFs computation in directions (scanning
method). \cref{sec:map} presents all algorithmic details for computing full CFs
maps (map method, full 2-point statistics for a given image). In
\cref{sec:verification} we verify the all computations based on known analytical
solutions. \cref{sec:efficiency} deals with computational efficiency of our
code, we compare computational times for 2D and 3D images of different sizes for
scanning and map methods as evaluated using either CPU or GPU. In
\cref{sec:examples} some application examples are provided, including CFs
evaluation for multi-phase materials. Finally, \cref{sec:summary} summarizes all
results and outlines possible future improvements.

\section{Overview of supported functions and modules.}
\label{sec:overview}
\verb+CorrelationFunctions.jl+ is capable to compute the following correlation
functions:
\begin{itemize}
\item Lineal-path function $L_2$.
\item Two-point function $S_2$.
\item Cluster function $C_2$.
\item Surface-surface function $F_{ss}$.
\item Surface-void function $F_{sv}$.
\item Pore size function $P$.
\item Chord length function $p$.
\item Phase cross-correlation function $\rho_{ij}$
\end{itemize}

These correlation functions are calculated using two different approaches (when
applicable). The first approach is implemented in the module \code{Directional}
and the second approach is in the module \code{Map}. The former approach (called
the scan approach) calculates correlation functions along specified directions
which can include axial and diagonal directions. The results of computations can
be provided to the user separately for each requested direction or averaged
across multiple directions. The latter approach (called the map approach)
calculates a ``correlation map'' which holds the values of a correlation
function in all directions. The scan and map approaches are described in
\cref{sec:scan} and \cref{sec:map} respectively.

@@

\section{The scan approach.}
\label{sec:scan}
In this approach, the input array is cut in parallel slices (``scanned'') along
each requested direction. Each slice is then analyzed independently as
one-dimensional system and the result is accumulated into \code{CorrelationData}
structure. In the following sections we provide the algorithms for computation
of the functions given in \cref{sec:overview}.

@@

\subsection{Lineal-path function}
Lineal-path function $L_2^{(i)}(\bm{r})$ is defined as the probability that a
line segment with the length $|\bm{r}|$ and paralel to $\bm{r}$ lies wholly in
the phase $i$ when randomly thrown into the input array. This function for a
one-dimensional array is computed using the algorithm described in
\apref{linpathalg}. A multidimensional arrays is firstly cut in slices and
\code{l2} function is calculated for each slice returning \code{success} and
\code{total} arrays. These arrays are then accumulated as described in
\cref{sec:scan} for final calculation of $L_2^{(i)}(\bm{r})$ function.

@@

\section{The map approach.}
\label{sec:map}

@@

\section{Verifying our algorithms}
\label{sec:verification}
To verify our package we use random datasets of overlapping disks (the number of
dimensions $n = 2$) or balls ($n = 3$) with radii $R$ and centers placed in
points generated by Poisson point process with a parameter $\lambda$. For these
datasets correlation functions can be given in closed-form expressions for one-,
two- and three-dimensional cases. For the sake of testing we obtain a collection
of disks (resp. balls) in a square (resp. cube) with the side $s$ using the
following algorithm:
\begin{algorithmic}[1]
  \label{testdata}
  \Procedure{testdata}{$s, n, R, \lambda$}
    \State $nballs \gets poisson(\lambda s^n)$ \Comment{$poisson(\lambda)$
      generates random numbers with Poisson distribution with parameter
      $\lambda$.}
    \State Generate $nballs$ random points which are uniformly distributes in
    a cube ($n = 3$) or square ($n = 2$) with a side $s$
    \State Draw balls (disks) of radius $R$ in those points. Store the result in
    an array $output$.
    \State \textbf{return} $output$
  \EndProcedure
\end{algorithmic}

@@

\section{Efficiency of our algorithms}
\label{sec:efficiency}
To measure efficiecy of our algorithms we use a machine with the following
configuration:
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 5 1600X
\item \textbf{Memory}: 32~Gb DDR4-2400 memory
\item \textbf{GPU}: Nvidia Tesla V100 with 32~Gb video memory
\end{itemize}

@@

\section{Examples and applications}
\label{sec:examples}
After the verification of our computational approach based on analytical
solution we can apply our algorithms to evaluate correlation functions for
different porous media images, including real XCT and SEM images.

@@

\section{Summary}
\label{sec:summary}
Text
